### 2015-07-26 science_writing/omsrs_writing.tex
adding a couple of tex that i want to work anywhere

\documentclass[a4paper, 11pt]{article}

\topmargin=-2.5cm
\textheight=10.5in
\oddsidemargin=-10pt
\textwidth=6.5in

\begin{document}

\begin{center}
\LARGE{Otago Medical School Research Society Science Writing Competition}\\
\end{center}

\section*{Why does my research matter?}
\noindent
\begin{itemize}
\item{obesity is a world-wide problem}
\item{cancer is a leading cause of death around the world}
\item{poor cancer prognosis due to subtypes}
\item{effect of chemotherapy varies between individuals}
\item{my research matters because it's looking for individual gene signatures that can be applied to personal medicine}
\end{itemize}

\section*{Explain the research:}
\noindent

\end{document}