### 2015-07-27 bioc192/intro.tex
added intro.tex

This book is intended to provide a basic overview of BIOC192 that will have more than enough details to tutor BIOC192 students.
Please note that to get the maximal use out of this book, and therefore how good you will become at being able to tutor someone in BIOC192, listening to the podcasts and adding your own notes is highly recommended.

Please DO NOT send this book around, especially to the HSFY students, unless you have asked the author (i.e. me) if you could do so.
All rights reserved.
No commercial reproduction, adaptation, distribution or transmission of this book is permitted without the written permission of \ldots me (Riku Takei).
And all the other legal crap people put on their books.

\vspace{5cm}

\noindent
Riku Takei \\
4$^{th}$ year Biochemistry MSc student\\