### 2015-07-29 bioc192/lect08.tex
finished lecture 8, onto lecture 9

\section{Introduction to Enzymes and Enzyme Engineering}

Enzymes (proteins) are able to accelerate and carry out chemical reactions that are most likely impossible to do under physiological/biological conditions.
Without these enzymes, we will not be able to synthesise chemicals required to survive.
It is also interesting that the enzymes have all evolutionarily developed itself to make difficult chemical reactions to faster, but within similar timescale.

Can we utilise the properties/characteristics of enzymes to make it do what we want it to do?
This is the key idea behind the researches being done in the area of enzyme engineering.
We can use enzyme engineering to make enzymes that are able to react with different substrates, or make it more efficient at doing the reaction.