### 2015-07-30 bioc192/tute.tex

\documentclass[a4paper, 12pt]{report}

\usepackage{chemmacros}
\usepackage{graphicx}
\usepackage[utf8]{inputenc}

\graphicspath{ {./images/topic1/} }

\newcommand{\HRule}{\rule {\linewidth}{0.5mm}}

\newcommand{\mychapter}[2]{
    \setcounter{chapter}{#1}
    \setcounter{section}{0}
    \chapter*{#2}
    \addcontentsline{toc}{chapter}{#2}
}

\begin{document}

\input{./title/title.tex}

\tableofcontents
\newpage

\mychapter{0}{Introduction}

\input{./intro.tex}

\mychapter{1}{Lecture 1}

\input{./lect01.tex}

\mychapter{2}{Lecture 2}

\input{./lect02.tex}

\mychapter{3}{Lecture 3}

\input{./lect03.tex}

\mychapter{4}{Lecture 4}

\input{./lect04.tex}

\mychapter{5}{Lecture 5}

\input{./lect05.tex}

\mychapter{6}{Lecture 6}

\input{./lect06.tex}

\mychapter{7}{Lecture 7}

\input{./lect07.tex}

\mychapter{8}{Lecture 8}

\input{./lect08.tex}

\mychapter{9}{Lecture 9}

\input{./lect09.tex}

\mychapter{10}{Lecture 10}

\input{./lect10.tex}

\mychapter{11}{Lecture 11}

\input{./lect11.tex}

\end{document}