### 2015-09-21 committee_report/committee_report.tex
added some specifications for committee report

\documentclass[12pt, a4paper]{article}

\usepackage[utf8]{inputenc}
\usepackage{natbib}
\usepackage{geometry}
\usepackage{fixltx2e}

\geometry{
    a4paper,
    left=2.5cm,
    right=2.5cm,
    top=2.5cm,
    bottom=2.5cm
}

\linespread{1.3}

\begin{document}

\section{50-word summary}
\section{introduction}
\section{methods (less than a page)}
\section{results}
\section{discussion}

results and discussion can be combined.

must not exceed 6 pages (with figures and references).

\citep{Hanahan2011}

%Referencing:
\bibliographystyle{BiocRefStyle}
\bibliography{../../../References/BibTeX/MSc}

\end{document}