### 2015-09-28 bioc192/tute.tex

\documentclass[a4paper, 12pt]{report}

\usepackage{chemmacros}
\usepackage{chemfig}
\usepackage{graphicx}
\usepackage[utf8]{inputenc}
\usepackage{siunitx}
\usepackage{geometry}

\geometry{
    a4paper,
    left=2.0cm,
    right=2.0cm,
    top=2.0cm,
    bottom=2.0cm
}

\graphicspath{ {./images/topic1/} {./images/topic2/} {./images/topic3/}{./images/topic5/}}

\newcommand{\HRule}{\rule {\linewidth}{0.5mm}}

\newcommand{\mychapter}[2]{
    \setcounter{chapter}{#1}
    \setcounter{section}{0}
    \chapter*{#2}
    \addcontentsline{toc}{chapter}{#2}
}

\NewChemState\bioGibbs[exponent=]{G}{\kilo\joule\per\mole}

\begin{document}

\input{./title/title.tex}

\tableofcontents
\newpage

\mychapter{0}{Introduction}

\input{./intro.tex}

\mychapter{1}{Lecture 1}

\input{./lect01.tex}

\mychapter{2}{Lecture 2}

\input{./lect02.tex}

\mychapter{3}{Lecture 3}

\input{./lect03.tex}

\mychapter{4}{Lecture 4}

\input{./lect04.tex}

\mychapter{5}{Lecture 5}

\input{./lect05.tex}

\mychapter{6}{Lecture 6}

\input{./lect06.tex}

\mychapter{7}{Lecture 7}

\input{./lect07.tex}

\mychapter{8}{Lecture 8}

\input{./lect08.tex}

\mychapter{9}{Lecture 9}

\input{./lect09.tex}

\mychapter{10}{Lecture 10}

\input{./lect10.tex}

\mychapter{11}{Lecture 11}

\input{./lect11.tex}

\mychapter{12}{Lecture 12}

\input{./lect12.tex}

\mychapter{13}{Lecture 13}

\input{./lect13.tex}

\mychapter{14}{Lecture 14}

\input{./lect14.tex}

\mychapter{15}{Lecture 15}

\input{./lect15.tex}

\mychapter{16}{Lecture 16}

\input{./lect16.tex}

\mychapter{17}{Lecture 17}

\input{./lect17.tex}

\mychapter{18}{Lecture 18}

\input{./lect18.tex}

\mychapter{19}{Lecture 19}

\input{./lect19.tex}

\mychapter{20}{Lecture 20}

\input{./lect20.tex}

\mychapter{21}{Lecture 21}

\input{./lect21.tex}

\mychapter{22}{Lecture 22}

\input{./lect22.tex}

\mychapter{23}{Lecture 23}

\input{./lect23.tex}

\mychapter{24}{Lecture 24}

\input{./lect24.tex}

\mychapter{25}{Lecture 25}

\input{./lect25.tex}

\mychapter{26}{Lecture 26}

\input{./lect26.tex}

\mychapter{27}{Lecture 27}

\input{./lect27.tex}

\mychapter{28}{Lecture 28}

\input{./lect28.tex}

\mychapter{29}{Lecture 29}

\input{./lect29.tex}

\mychapter{30}{Lecture 30}

\input{./lect30.tex}

\mychapter{31}{Lecture 31}

\input{./lect31.tex}

\mychapter{32}{Lecture 32}

\input{./lect32.tex}

\mychapter{33}{Lecture 33}

\input{./lect33.tex}

\mychapter{34}{Lecture 34}

\input{./lect34.tex}

\mychapter{35}{Lecture 35}

\input{./lect35.tex}

\end{document}