### 2015-10-03 bioc192/tute.tex

\documentclass[a4paper, 12pt]{report}

\usepackage{chemmacros}
\usepackage{chemfig}
\usepackage{graphicx}
\usepackage[utf8]{inputenc}
\usepackage{siunitx}
\usepackage{geometry}
\usepackage{newclude}

\geometry{
    a4paper,
    left=2.0cm,
    right=2.0cm,
    top=2.0cm,
    bottom=2.0cm
}

\graphicspath{ {./images/topic1/} {./images/topic2/} {./images/topic3/}{./images/topic5/}{./images/topic6/}}

\newcommand{\HRule}{\rule {\linewidth}{0.5mm}}

\newcommand{\mychapter}[2]{
    \setcounter{chapter}{#1}
    \setcounter{section}{0}
    \chapter*{#2}
    \addcontentsline{toc}{chapter}{#2}
}

\NewChemState\bioGibbs[exponent=]{G}{\kilo\joule\per\mole}

\includeonly{./lect36,./lect37,./lect38,./lect39}

\begin{document}

\include*{./title/title}

\tableofcontents
\newpage

\mychapter{0}{Introduction}

\include*{./intro}

\mychapter{1}{Lecture 1}

\include*{./lect01}

\mychapter{2}{Lecture 2}

\include*{./lect02}

\mychapter{3}{Lecture 3}

\include*{./lect03}

\mychapter{4}{Lecture 4}

\include*{./lect04}

\mychapter{5}{Lecture 5}

\include*{./lect05}

\mychapter{6}{Lecture 6}

\include*{./lect06}

\mychapter{7}{Lecture 7}

\include*{./lect07}

\mychapter{8}{Lecture 8}

\include*{./lect08}

\mychapter{9}{Lecture 9}

\include*{./lect09}

\mychapter{10}{Lecture 10}

\include*{./lect10}

\mychapter{11}{Lecture 11}

\include*{./lect11}

\mychapter{12}{Lecture 12}

\include*{./lect12}

\mychapter{13}{Lecture 13}

\include*{./lect13}

\mychapter{14}{Lecture 14}

\include*{./lect14}

\mychapter{15}{Lecture 15}

\include*{./lect15}

\mychapter{16}{Lecture 16}

\include*{./lect16}

\mychapter{17}{Lecture 17}

\include*{./lect17}

\mychapter{18}{Lecture 18}

\include*{./lect18}

\mychapter{19}{Lecture 19}

\include*{./lect19}

\mychapter{20}{Lecture 20}

\include*{./lect20}

\mychapter{21}{Lecture 21}

\include*{./lect21}

\mychapter{22}{Lecture 22}

\include*{./lect22}

\mychapter{23}{Lecture 23}

\include*{./lect23}

\mychapter{24}{Lecture 24}

\include*{./lect24}

\mychapter{25}{Lecture 25}

\include*{./lect25}

\mychapter{26}{Lecture 26}

\include*{./lect26}

\mychapter{27}{Lecture 27}

\include*{./lect27}

\mychapter{28}{Lecture 28}

\include*{./lect28}

\mychapter{29}{Lecture 29}

\include*{./lect29}

\mychapter{30}{Lecture 30}

\include*{./lect30}

\mychapter{31}{Lecture 31}

\include*{./lect31}

\mychapter{32}{Lecture 32}

\include*{./lect32}

\mychapter{33}{Lecture 33}

\include*{./lect33}

\mychapter{34}{Lecture 34}

\include*{./lect34}

\mychapter{35}{Lecture 35}

\include*{./lect35}

\mychapter{36}{Lecture 36}

\include*{./lect36}

\mychapter{37}{Lecture 37}

\include*{./lect37}

\mychapter{38}{Lecture 38}

\include*{./lect38}

\mychapter{39}{Lecture 39}

\include*{./lect39}

\end{document}