### 2016-03-10 thesis/thesis.tex
starting my thesis - just the backbone structure for now

\documentclass[a4paper, 12pt, oneside]{report}

% Preamble:
\usepackage[T1]{fontenc}       % package for specifying output font encoding
\usepackage[utf8]{inputenc}    % package for specifying input font encoding
\usepackage{graphicx}          % package for graphics handling (i.e. figure images)
\usepackage{mathptmx}          % package for Times New Roman font
\usepackage{setspace}          % package for setting line space

% If you want to compile specific files, uncomment the line below and add the filenames in the {} :
%\includeonly{}

\begin{document}

\include{./title}

\pagenumbering{roman} % set page numbering to roman numeral

\include{./abstract}

\include{./acknowledgement}

\tableofcontents

\addcontentsline{toc}{section}{\contentsname}

\clearpage

\addcontentsline{toc}{section}{\listfigurename}

\listoffigures

\clearpage

\addcontentsline{toc}{section}{\listtablename}

\listoftables

% May need to include abbreviations and glossary here, before any chapter begins

\clearpage

\pagenumbering{arabic} % revert the page numbering to normal numbering

\include{./intro}

\include{./methods}

\include{./results}

\include{./discussion}

\end{document}