### 2016-03-10 thesis/title.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Title page for my thesis - code is pretty self explanatory
% Required packages: setspace
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{titlepage}

\centering

\rule[2.0mm]{\textwidth}{0.5mm}

\begin{doublespace}
    {\Huge \scshape Body Mass Index and Pathway Dysregulation in Cancer}
\end{doublespace}

\rule[2.0mm]{\textwidth}{0.5mm}

\vspace{2.0mm}
{\Large \scshape Riku Takei}

\vfill

{\normalsize \scshape A Thesis Submitted For The Degree Of\\}
{\Large \scshape Master Of Science\\}
\vspace{2.0mm}
{\normalsize \scshape At The University Of Otago, Dunedin, New Zealand \\}

\vspace{10.0mm}
{\normalsize \today}

\end{titlepage}