### 2016-03-30 tips_and_notes/vim/chapter3/command.tex
onto Ex mode

\section{Command Line (Ex) Mode}

This mode is essentially an extension of normal mode, but provides vim with more functionality and super effective at editing the whole document (if used correctly).
With normal mode by itself, you can only edit so much with the given operator and motion keys.
You can increase and specify the range with visual modes, but this will only affect locally, within the selection.
If you wanted to edit the whole document, you can use \verb|G| as  the motion, but the operator keys may not be specific enough to do the task you want it to do.

Ex mode provides the minor adjustments and specificity required by you to carry out the exact task you want vim to do for the whole document, or just part of it.