### 2016-03-30 tips_and_notes/vim/chapter3/visual.tex

\section{Visual Mode}

Visual mode is very similar to normal mode, but it allows you to define text selection using your cursor position(s).
You can enter visual mode by pressing \verb|v| in normal mode.
To enter a line-wise visual mode, press \verb|V|, and for a block-wise visual mode, press \verb|<ctrl-v>|.

To reselect the previous selection, press \verb|gV|.
If you want to swap the cursor position from one end of a selection to another, press \verb|o| while in insert mode.

If you want to carry out a (ex) command on the selection, press \verb|:| while in insert mode with the text selected (or just simply type the operator command if it's not complicated).
This will drop you into ex-command line prompt, where you can type in whatever you want to do with the selection.