### 2016-04-01 tips_and_notes/vim/chapter4/files.tex

\chapter{Files, Windows, and Tabs}

\section{Files}