### 2016-04-01 tips_and_notes/vim/chapter4/window.tex
finished chapter 4

\section{Windows}

Vim allows us to split the window with same or different buffers for editing.
An alternative to this would be to use tmux.
You can use windows to look at different parts of the document, so that you can have a window open to refer to the text elsewhere of the document, and a window open for editing the document.

\subsection{Split and navigate through windows}

\begin{tabular}{c|l}
\verb|<ctrl-w>s|              & split window horizontally\\
\verb|<ctrl-w>h|              & split window vertically\\
\verb|:sp {file}|             & split window horizontally and load file\\
\verb|:vsp {file}|            & split window vertically and load file\\
\verb|<ctrl-w><ctrl-w>|       & cycle through open windows\\
\verb|<ctrl-w><ctrl-h/j/k/l>| & focus the window to left/down/up/right \\
                              & window\\
\verb|:cl(ose)|               & close the active window\\
\verb|:on(ly)|                & close all windows except current window\\
\verb|<ctrl-w>=|              & equalize width and height of all windows\\
\verb|<ctrl-w>_|              & maximize height of the active window\\
\verb=<ctrl-w>|=              & maximize width of the active window\\
\verb|[N]<ctrl-w>_|           & set active window height to N rows\\
\verb=[N]<ctrl-w>|=           & set active window width to N columns\\
\end{tabular}