### 2016-04-04 tips_and_notes/vim/chapter4/tab.tex

\section{Tabs}

Tabs in vim can hold a collection of windows (like a workspace in  linux).
The \verb|:lcd {path}| command sets the working directory for the current window (not the tab).
This means that you can assign different working directory to different windows/tabs (i.e. separate workspace).
\verb|:tabedit {filename}| will open a new tab with filename.
If filename is empty, tab is created with an empty buffer.
\verb|<ctrl-w>T| will move the current window into a new tab page.
\verb|:tabclose| will close the tab page, even if it has got multiple windows open.
\verb|:tabonly| will close all other tabs except for the current tab.
Finally, \verb|gt/gT| will ``go to'' the next/previous tab.
You can specify the number of tabs you want to jump past (e.g. 3gt).