### 2016-04-05 tips_and_notes/vim/chapter5/efficient.tex
wrote navigation summary

\section{Efficient navigation}

\section{Navigation Summary}

\begin{center}
\begin{tikzpicture}[node distance = 1.5cm]
    \node (o) at (0,0) {};
    \node [left=of o] (h) {h/b/B/ge/gE/0/T/F/?};
    \node [right=of o] (l) {l/w/W/e/E/\$/t/f/\verb|/|};
    \node [above=of o] (k) {k/\verb|<ctrl-b>|};
    \node [below=of o] (j) {j/\verb|<ctl-f>|};
    \draw [<->, very thick] (h) edge (l) (k) edge (j);
\end{tikzpicture}
\end{center}