### 2016-04-07 tips_and_notes/vim/chapter7/macros.tex
started on chapter 7

\chapter{Macros}

From the Practical Vim: \textit{``Like the game of Othello, Vim's macros takes a minute to learn and a lifetime to master.''}

\section{Record and execute a macro}

The \verb|q| key acts as both the start and stop command for recording a macro\verb|q| key acts as both the start an stop command for recording a macro,
\verb|q{register}| starts recording our keystrokes to be used as a macro.
Press \verb|q| again to stop recording the keystrokes.
Note that anything stored in the register will be overwritten after you have recorded the macro to the register.

To execute