### 2016-04-08 tips_and_notes/vim/chapter8/patterns.tex
started chapter 8

\chapter{Search and pattern matching}

\section{Basic searching}

As discussed before, \verb|/| searches forward and \verb|?| searches backward for a pattern, and \verb|n/N| will jump to the next/previous search pattern.
Repeat the search by leaving the search field blank and press \verb|<CR>|
Note that \verb|n| goes forward for \verb|/| search but ``backward'' (forward in the backward direction) for the \verb|?| search.
To switch the direction, just use \verb|/| or \verb|?| with no search pattern -- this will re-use the previous pattern to search forward/backward.
You can go through your search history by using the \verb|<up>| key in the search prompt (or use \verb|q/|).

\subsection{Mute search highlighting}

\verb|:noh(lsearch)| will mute the highlighting for current search, but will enable it for the next search.
Either use this as is (\verb|:noh|) or map it to a useful shortcut.

\section{Pattern matching in search}

There are various tricks and tips you can add to your search pattern to hone in  on the target text that you are actually interested in.

\subsection{Case sensitivity in search}

You can set the case sensitivity globally by tweaking the \verb|:ignorecase| setting, but you may want to set the case sensitivity differently, depending on your needs.
The \verb|\c| and \verb|\C| items in your search pattern will allow you to do this.