### 2016-04-14 thesis/thesis.tex
organised the directory structure

\documentclass[a4paper, 12pt, oneside]{report}

% Preamble:
\usepackage[T1]{fontenc}         % package for specifying output font encoding
\usepackage[utf8]{inputenc}      % package for specifying input font encoding
\usepackage{graphicx}            % package for graphics handling (i.e. figure images)
\usepackage{mathptmx}            % package for Times New Roman font
\usepackage{setspace}            % package for setting line space
\usepackage[hidelinks]{hyperref} % package for hyperlinking references - hidelinks option removes the ugly boxes around the huperlinks

% Set up the hyperlinks - hange the colour of the links to blue:
\hypersetup{colorlinks=true,linkcolor=blue}

% If you want to compile specific files, uncomment the line below and add the filenames in the {} :
%\includeonly{}

\begin{document}

\include{./misc/title}

\pagenumbering{roman} % set page numbering to roman numeral

\include{./misc/abstract}

\include{./misc/acknowledgement}

\tableofcontents

\addcontentsline{toc}{section}{\contentsname}

\clearpage

\addcontentsline{toc}{section}{\listfigurename}

\listoffigures

\clearpage

\addcontentsline{toc}{section}{\listtablename}

\listoftables

% May need to include abbreviations and glossary here, before any chapter begins

\clearpage

\pagenumbering{arabic} % revert the page numbering to normal numbering

\include{./intro/intro}

\include{./methods/methods}

\include{./results/results}

\include{./discussion/discussion}

\end{document}