### 2016-04-15 thesis/intro/intro.tex
added a bit of plan in my thesis; added few more file types in .gitignore; added reference style file

\chapter{Introduction}

\section{Obesity}

General intro to obesity\\
Implications of obesity -- disease associated with it and/or cancer\\

\section{Cancer}

What is cancer\\

\subsection{Hallmarks of Cancer}

What is it caused by (11 hallmarks of cancer)\\

\subsection{Relationship between obesity}

Why is it important in obesity context\\

\section{Obesity Associated Genetic Signatures}

Literature review, focussing on \citet{Creighton2012} and \citet{Fuentes-Mattei2014}\\

\section{Aim Of The Project}