### 2016-04-15 thesis/methods/methods.tex

\chapter{Methods}

\section{Download Data from publicly available source}

\subsection{The Cancer Genome Atlas}

\subsection{International Cancer Genome Consortium}

\section{Validation Of Methodology}

\subsection{Normalisation of \citet{Creighton2012} data}

where was the original data from?\\
(rma normalisation before any analysis)

\subsection{Differential Gene Expression Analysis}

(mention WHO classification of obese/overweight/lean)\\
(also mention limma package)

\subsection{Singular Value Decomposition}

\section{Applying metagene to other cancer types}

\subsection{Transformation matrix}

\section{Pathway Enrichment Analysis}