### 2016-04-15 thesis/results/results.tex

\chapter{Results from \citet{Creighton2012} metagene}

\section{SVD}