### 2016-04-16 thesis/methods/methods.tex
started working on methods section

\chapter{Methods}

\section{Body Mass Index (BMI)}

\subsection{BMI Calculation}

BMI was calculated from the height and weight data of the samples using the following equation:\\
\begin{equation}
    BMI = \frac{weight (kg)}{Height^2(m^2)}
\end{equation}

\subsection{BMI Classification}

Samples were classified based on the WHO definition:
\begin{table}[h]
    \caption{WHO defined BMI classification}\\
    \label{tab:who-bmi-class}
    \begin{center}
        \begin{tabular}{|c|c|}
            \hline
            \bfseries {Classification} & \bfseries {BMI}\\
            \hline
            Underweight & \textless 20.0\\
            Normal weight/lean & 20.0$\sim$24.9\\
            Overweight & 25.0$\sim$29.9\\
            Obese & \textgreater 30.0\\
            \hline
        \end{tabular}
    \end{center}
\end{table}

\section{Download Data from publicly available source}

Data related to multiple different cancer types were downloaded from free and publicly available source such as The Cancer Genome Atlas and the International Cancer Genome Consortium.

\subsection{The Cancer Genome Atlas (TCGA)}

The clinical data for all cancer types (33 types in total) were downloaded from TCGA database and were checked for the height and weight data for each sample.
Any cancer type with no weight and/or height data of the samples were excluded from the project, as no BMI information can be obtained without these data.
This gave a total of eight cancer types (**insert all the names of cancer types here) to be used for further analyses.

\subsection{International Cancer Genome Consortium (ICGC)}

RNA-seq data for the eight cancer types with sample BMI information were downloaded from the ICGC database.

\section{Validation of methodology used by \citet{Creighton2012}}

\subsection{Raw data from \citet{Creighton2012} paper}

The raw microarray gene expression data files of all the samples used by \citet{Creighton2012} were downloaded from the Gene Expression Omnibus (GEO) database.
Raw data were imported into R/RStudio for further analyses.

\subsection{Normalisation of \citet{Creighton2012} data}

The raw microarray data for all 103 samples in the study were normalised using Robust Multi Array (RMA) method from the \textit{affy} package in R.

\subsection{Differential gene expression analysis}

(mention WHO classification of obese/overweight/lean)\\
(also mention limma package)

\subsection{Singular Value Decomposition}

\section{Applying metagene to other cancer types}

\subsection{Transformation matrix}

\section{Pathway Enrichment Analysis}