### 2016-04-18 thesis/intro/intro.tex
tmp commit

\chapter{Introduction}
\label{ch:intro}

\section{Obesity}
\label{sec:obesity}

General intro to obesity.
Implications of obesity -- disease associated with it and/or cancer.

\section{Cancer}
\label{sec:cancer}

What is cancer.

\subsection{Hallmarks of Cancer}
\label{subsec:cancerhallmarks}

What is it caused by (11 hallmarks of cancer).

\subsection{Relationship between obesity}
\label{subsec:obsbackground}

Why is it important in obesity context.

\section{Obesity associated genetic signatures}
\label{sec:obsgene}

Literature review, focussing on \citet{Creighton2012} and \citet{Fuentes-Mattei2014}.

\section{Aim of the project}
\label{sec:aim}

This research project aims to determine whether gene expression signatures exist that are specific to obesity across multiple different cancer types, and to investigate whether there are any common pathways being dysregulated in cancers based on these genetic signatures.
Better understanding of the pathways being dysregulated in cancer cells in obese patients may lead to improved clinical decisions, and could thus contribute towards personalised treatment in clinical settings in the future.