### 2016-04-22 summary_report/summary.tex
started writing up a summary in prep to committee meeting; some work done on the thesis

\documentclass[a4paper, 11pt]{article}

\usepackage[utf8]{inputenc}
\begin{document}

\section*{Summary so far}

\subsection*{Creighton data}

Creighton et al. found obesity-associated genetic signatures that are specific to obese samples, but not in the lean or overweight samples, in a study using breast tumour samples from patients.

\end{document}