### 2016-05-09 bioc192/intro.tex
editing bioc notes

This book is intended to provide a ``basic'' overview of BIOC192 that will have enough details to tutor BIOC192 students.
It is highly recommended to listen to the podcasts and add your own notes to this book to get the maximal use out of this book.

This book is organised in such a way that the contents/topics of each lecture are more in a logical order (in my opinion), rather than in the order the lecturer presents the material.
I feel like some of the lecturers skip some details, go back to it later on, and go back and forth every now and then.
This  is quite understandable as they only have an hour to go through a certain aspect of the overall concept, and hence after a block of lectures, you usually understand most of what they were saying in the previous lectures.
Since this book is not time limited, I have organised it into  a ``logical order'' as best as I can.

Please DO NOT send this book around, especially to the HSFY students, unless you have asked the author (i.e. me) if you could do so.
All rights reserved.
No commercial reproduction, adaptation, distribution or transmission (and any other legal crap people put on their books) of this book is permitted without the written permission of the author.

\vspace{5cm}

\noindent
Riku Takei \\
5$^{th}$ year Biochemistry MSc student\\