### 2016-09-12 presentation/beamer_theme/ctcrSep2016.tex
draft ctcr presentation

% \documentclass[notes]{beamer}
% \documentclass[notes=hide]{beamer}
% \documentclass[notes=only]{beamer}
\documentclass[handout]{beamer}

% \usepackage{pgfpages}
% \pgfpagesuselayout{2 on 1}[a4paper,border shrink=3mm]

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{xcolor,colortbl}

\definecolor{LightCyan}{rgb}{0.88,1,1}

\usetheme{riku}

\def\logo{
	{\includegraphics[width=20mm]{logo}}
}

\title[BMI and Cancer]{Body Mass Index and Pathway Dysregulation in Cancer}
\subtitle{CTCR Talk}
\author[Riku Takei]{Riku Takei (MSc)\\Supervisor: Assoc. Prof. Mik Black}
\institute{Department of Biochemistry, University of Otago}

\begin{document}

{
	\setbeamertemplate{footline}{}
	\begin{frame}[noframenumbering]
		\titlepage
	\end{frame}
}

\begin{frame}
	\frametitle{Introduction}
	\begin{center}
		\includegraphics[scale=0.35]{./images/obesityworldmap}
	\end{center}
\end{frame}

\begin{frame}
	\frametitle{Introduction}
	\begin{center}
		\includegraphics[width=0.48\paperwidth]{./images/ob1999}\\
		1999\\
		\includegraphics[width=0.48\paperwidth]{./images/ob2014}\\
		2014
	\end{center}
\end{frame}

\begin{frame}
	\frametitle{Body Mass Index and Cancer}
	\begin{columns}
		\column{0.6\textwidth}
		{\footnotesize
			Cancers are caused by dysregulation in various pathways/mechanisms that allow them to survive and proliferate
			\begin{itemize}
				\item Commonly know as the Hallmarks of Cancer
			\end{itemize}
			Many reports on obesity being a major risk for cancer and poor prognosis
		}
		\column{0.5\textwidth}
		\includegraphics[width=0.45\paperwidth]{./images/cancerhallmark}
	\end{columns}
\end{frame}

\begin{frame}
	\frametitle{Aims}
	\begin{enumerate}
		\item Investigate the association between patient BMI and tumour molecular characteristics
		\item Determine whether pathway commonalities exist across tumours that arise from other tissues
	\end{enumerate}
\end{frame}

\begin{frame}
	\frametitle{Obesity-associated Genetic Signatures}
	Creighton \textit{et al.} (2012)
	\begin{itemize}
		\item Identified obesity-associated genetic signatures from the tumours of breast cancer patients
		\item Showed that some compounds were able to specifically regulate the expression of these genetic signatures (potential for targeted therapy)
	\end{itemize}
\end{frame}

\begin{frame}{{\normalsize Obesity-associated genes from Creighton \textit{et al.} (2012) }}
	103 samples used in the study, 799 obesity-associated gene probes (695 unique genes) identified\\
	{\footnotesize
		\begin{itemize}
			\item Majority of these genes were downregulated in obese patients
		\end{itemize}
	}
	\begin{center}
		\includegraphics[width=0.6\paperwidth]{./images/creighton}
	\end{center}
\end{frame}

\begin{frame}{Obesity-associated genes from Fuentes-Mattei \textit{et al.}}
	\begin{columns}
		\column{0.5\textwidth}
		Fuentes-Mattei \textit{et al.} (2014)
		\begin{itemize}
			\item 137 ER+ breast cancer patients in the study, 130 obesity-associated genes identified
			\item Direct evidence of how obesity/adiposity contributes to tumour growth (using mouse model)
		\end{itemize}
		\column{0.5\textwidth}
		\includegraphics[width=0.45\paperwidth]{./images/fmcircos}
	\end{columns}
\end{frame}

\begin{frame}
	\frametitle{Methods}
	Downloaded RNA-seq data from 8 different cancer types from ICGC
	\vskip2ex
	Microarray data from Creighton's paper
	\vskip2ex
	Microarray data from Fuentes-Mattei's paper
	\vskip2ex
	Microarray data from Cris Print
	\vskip2ex
	\includegraphics[width=0.4\textwidth]{./images/tcga}
	\hfill
	\includegraphics[width=0.4\textwidth]{./images/icgc}
\end{frame}

\begin{frame}{Validation of Methodology -- Creighton \textit{et al.} (2012)}
	103 samples used in the study, 799 obesity-associated gene probes (695 unique genes) identified\\
	\vskip2ex
	Metagene was obtained from SVD using the genes that were common in both Creighton \textit{et al.} and other cancer data sets\\
	\begin{center}
		\includegraphics[width=0.6\paperwidth]{./images/creighton}
	\end{center}
\end{frame}

\begin{frame}{Singular Value Decomposition (SVD) recap}
	For each sample in the data set, you need to assign a ``score'' that represents the overall gene expression of the gene set of interest
	\begin{itemize}
		\item i.e. a score that represent the 799 obesity-associated gene probes identified by Creighton \textit{et al.}
	\end{itemize}
	\vskip2ex
	You can do this by applying SVD to the data set of interest and get a metagene out from this
\end{frame}

\begin{frame}{\normalsize Creighton \textit{et al.} metagene correlates with Gene Expression}
	\begin{center}
		\includegraphics[width=0.45\paperwidth]{./images/crge1}
	\end{center}
\end{frame}

\begin{frame}{Creighton \textit{et al.} metagene correlates with BMI}
	\hspace{-4.0ex}
	\includegraphics[width=0.45\paperwidth]{./images/crmetaclin0}
	\includegraphics[width=0.45\paperwidth]{./images/crmetaclin1}
\end{frame}

\begin{frame}{Transformation Matrix}
	Singular Value Decomposition of a dataset X gives:
	\begin{equation*}
		X = UDV'
	\end{equation*}
	Rearranging this equation gives:
	\begin{equation*}
		V' = D^{-1}U'X
	\end{equation*}
	Where $D^{-1}U'$ is the transformation matrix and $V'$ is the metagene for the data set X
	{\footnotesize
		\begin{itemize}
			\item This means that the transformation matrix can be applied to other data sets to obtain the metagene for those data sets
		\end{itemize}
	}
\end{frame}

\begin{frame}{Transformation Matrix}
	\begin{center}
		\includegraphics[height=0.7\paperheight,width=0.40\paperwidth]{./images/tmp/transmatexp1}
	\end{center}
\end{frame}

\begin{frame}{Transformation Matrix}
	\begin{center}
		\includegraphics[width=0.70\paperwidth]{./images/tmp/transmatexp2}
	\end{center}
\end{frame}

\begin{frame}{Metagene in ICGC data}
	\includegraphics[width=0.23\paperwidth]{./images/crtcga2}
	\includegraphics[width=0.23\paperwidth]{./images/crtcga7}
	\includegraphics[width=0.23\paperwidth]{./images/crtcga12}
	\includegraphics[width=0.23\paperwidth]{./images/crtcga17}\\
	\includegraphics[width=0.23\paperwidth]{./images/crtcga22}
	\includegraphics[width=0.23\paperwidth]{./images/crtcga27}
	\includegraphics[width=0.23\paperwidth]{./images/crtcga32}
	\includegraphics[width=0.23\paperwidth]{./images/crtcga37}
\end{frame}

\begin{frame}{Does the metagene correlate with BMI?}
	\includegraphics[width=0.23\paperwidth]{./images/crtcga3}
	\includegraphics[width=0.23\paperwidth]{./images/crtcga8}
	\includegraphics[width=0.23\paperwidth]{./images/crtcga13}
	\includegraphics[width=0.23\paperwidth]{./images/crtcga18}\\
	\includegraphics[width=0.23\paperwidth]{./images/crtcga23}
	\includegraphics[width=0.23\paperwidth]{./images/crtcga28}
	\includegraphics[width=0.23\paperwidth]{./images/crtcga33}
	\includegraphics[width=0.23\paperwidth]{./images/crtcga38}
\end{frame}

\begin{frame}{{\normalsize Obesity-associated genes from Fuentes-Mattei \textit{et al.}}}
	\begin{columns}
		\column{0.5\textwidth}
		137 ER+ breast cancer samples in the study, 130 unique genes identified\\
		\vskip2ex
		Data were normalised  with RMA normalisation method\\
		\vskip2ex
		\column{0.5\textwidth}
		\includegraphics[width=0.45\paperwidth]{./images/fmcircos}
	\end{columns}
\end{frame}

\begin{frame}{FM metagene in Creighton \textit{et al.} data}
	\begin{center}
		\includegraphics[width=0.45\paperwidth]{./images/fmmetacr3}
	\end{center}
\end{frame}

\begin{frame}{FM metagene in Creighton \textit{et al.} data}
	\includegraphics[width=0.45\paperwidth]{./images/fmmetacr4}
	\includegraphics[width=0.45\paperwidth]{./images/fmmetacr5}
\end{frame}

\begin{frame}{FM metagene in other cancer types}
	\includegraphics[width=0.23\paperwidth]{./images/fmmetaheat2}
	\includegraphics[width=0.23\paperwidth]{./images/fmmetaheat7}
	\includegraphics[width=0.23\paperwidth]{./images/fmmetaheat12}
	\includegraphics[width=0.23\paperwidth]{./images/fmmetaheat17}\\
	\includegraphics[width=0.23\paperwidth]{./images/fmmetaheat22}
	\includegraphics[width=0.23\paperwidth]{./images/fmmetaheat27}
	\includegraphics[width=0.23\paperwidth]{./images/fmmetaheat32}
	\includegraphics[width=0.23\paperwidth]{./images/fmmetaheat37}
\end{frame}

\begin{frame}{FM metagene and BMI}
	\includegraphics[width=0.23\paperwidth]{./images/fmmetabmi3}
	\includegraphics[width=0.23\paperwidth]{./images/fmmetabmi8}
	\includegraphics[width=0.23\paperwidth]{./images/fmmetabmi13}
	\includegraphics[width=0.23\paperwidth]{./images/fmmetabmi18}\\
	\includegraphics[width=0.23\paperwidth]{./images/fmmetabmi23}
	\includegraphics[width=0.23\paperwidth]{./images/fmmetabmi28}
	\includegraphics[width=0.23\paperwidth]{./images/fmmetabmi33}
	\includegraphics[width=0.23\paperwidth]{./images/fmmetabmi38}
\end{frame}

\begin{frame}{Gene Expression Analysis}
	Maybe the obesity-associated genes identified by either group were not related to obesity, but related to a different clinical variable
	\vskip2ex
	Are the differentially expressed genes between the obese and non-obese groups (after correcting for other clinical variables) similar to the obesity-associated genes?
	{\footnotesize
		\begin{itemize}
			\item How do these DEGs compare with the Creighton's or Fuentes-Mattei's obesity-associated genes?
		\end{itemize}
	}
\end{frame}

\begin{frame}{Creighton \textit{et al.} data revisited}
	Gene expression analysis was carried out on the Creighton \textit{et al.} data:
	\begin{itemize}
		\item Original Creighton data
		\item Caucasian-only Creighton data
	\end{itemize}
	\vskip2ex
	Gene expression analysis was repeated with the residual version of the above data
	\vskip2ex
	Each ``obesity-associated genes'' were checked for any overlaps with the original Creighton metagene
\end{frame}

\begin{frame}{Completely remove the effect of ethnicity}
	African American patients are more likely to have higher BMI than Cauacasian patients
	\vskip2ex
	Removed the effect of ethnicity by removing the African American patients' data (total of 77 patient data to work with)
	{\footnotesize
		\begin{itemize}
			\item All of the other clinical variables were adjusted as before
		\end{itemize}
	}
\end{frame}

\begin{frame}{DEGs in residual data}
	\begin{center}
		\includegraphics[width=0.42\paperwidth]{./images/venn0}
		\includegraphics[width=0.45\paperwidth]{./images/venn1}
	\end{center}
\end{frame}

\begin{frame}{Metagene Analysis}
	All of the gene probe sets I have found (8 in total) were used to create metagenes as before
	\begin{itemize}
		\item Checked in Creighton \textit{et al.} data
		\item Transferred into other cancer data sets
	\end{itemize}
\end{frame}

\begin{frame}{Initial results -- Checking in Creighton \textit{et al.} data}
	\includegraphics[width=0.23\paperwidth]{./images/degmetacr2}
	\includegraphics[width=0.23\paperwidth]{./images/degmetacr7}
	\includegraphics[width=0.23\paperwidth]{./images/degmetacr12}
	\includegraphics[width=0.23\paperwidth]{./images/degmetacr17}\\
	\includegraphics[width=0.23\paperwidth]{./images/degmetacr22}
	\includegraphics[width=0.23\paperwidth]{./images/degmetacr27}
	\includegraphics[width=0.23\paperwidth]{./images/degmetacr32}
	\includegraphics[width=0.23\paperwidth]{./images/degmetacr37}
\end{frame}

\begin{frame}{Initial results -- Checking in Creighton \textit{et al.} data}
	\includegraphics[width=0.23\paperwidth]{./images/degmetacr3}
	\includegraphics[width=0.23\paperwidth]{./images/degmetacr8}
	\includegraphics[width=0.23\paperwidth]{./images/degmetacr13}
	\includegraphics[width=0.23\paperwidth]{./images/degmetacr18}\\
	\includegraphics[width=0.23\paperwidth]{./images/degmetacr23}
	\includegraphics[width=0.23\paperwidth]{./images/degmetacr28}
	\includegraphics[width=0.23\paperwidth]{./images/degmetacr33}
	\includegraphics[width=0.23\paperwidth]{./images/degmetacr38}
\end{frame}

\begin{frame}{Do the metagenes correlate with GE in ICGC data?}
	\includegraphics[width=0.23\paperwidth]{./images/rawobsmeta2}
	\includegraphics[width=0.23\paperwidth]{./images/resobsmeta2}
	\includegraphics[width=0.23\paperwidth]{./images/caobsmeta2}
	\includegraphics[width=0.23\paperwidth]{./images/caresobsmeta2}\\
	\includegraphics[width=0.23\paperwidth]{./images/crolmeta2}
	\includegraphics[width=0.23\paperwidth]{./images/rescrolmeta2}
	\includegraphics[width=0.23\paperwidth]{./images/cacrolmeta2}
	\includegraphics[width=0.23\paperwidth]{./images/carescrolmeta2}
\end{frame}

\begin{frame}{Do the metagenes correlate with BMI?}
	\includegraphics[width=0.23\paperwidth]{./images/rawobsclin3}
	\includegraphics[width=0.23\paperwidth]{./images/resobsclin3}
	\includegraphics[width=0.23\paperwidth]{./images/caobsclin3}
	\includegraphics[width=0.23\paperwidth]{./images/caresobsclin3}\\
	\includegraphics[width=0.23\paperwidth]{./images/crolclin3}
	\includegraphics[width=0.23\paperwidth]{./images/rescrolclin3}
	\includegraphics[width=0.23\paperwidth]{./images/cacrolclin3}
	\includegraphics[width=0.23\paperwidth]{./images/carescrolclin3}
\end{frame}

% \begin{frame}
% 	Repeated this with Cris' data, but same results as in Creighton and ICGC data
% \end{frame}

% \begin{frame}{Do the metagenes correlate with each other?}
% 	\begin{center}
% 		\includegraphics[width=0.45\paperwidth]{./images/cor1}
% 	\end{center}
% \end{frame}

% \begin{frame}{Common genes across multiple cancer types}
% 	DEGs between the obese and non-obese groups were identified for each cancer types
% 	\vskip2ex
% 	For each cancer types, samples were randomly split into two groups (same proportion as obese vs. non-obese samples)
% 	\vskip2ex
% 	{\footnotesize
% 		\begin{tabular}{ccccccccc}
% 			& \multicolumn{8}{c}{No. of cancer types the gene overlapped}\\
% 			\rule{0pt}{2.25ex}                    & 1    & 2    & 3   & 4  & 5 & 6 & 7 & 8 \\
% 			\hline
% 			\rule{0pt}{2.25ex}Obese vs. Non-obese & 6865 & 1907 & 289  & 26   & 1 & 0 & 0 & 0 \\
% 			Randomly split groups                 & 6202 & 1488 & 198  & 18   & 1 & 0 & 0 & 0 \\
% 			No. of genes found                    & 5012 & 870  & 89   & 6    & 0 & 0 & 0 & 0 \\
% 			on average                            &      &      &      &      &   &   &   & \\
% 			Empirical FDR                         & 0.73 & 0.46 & 0.31 & 0.23 & 0 & 0 & 0 & 0 \\
% 		\end{tabular}
% 	}
% \end{frame}

% \begin{frame}{Pathway Enrichment Analysis}
% 	Maybe each cancer type have a different set of genes activated
% 	\vskip2ex
% 	Are there any specific pathways enriched in any specific cancer type?
% 	\vskip2ex
% 	If so, is it common across other cancer types?
% 	\vskip2ex
% 	How relevant is the pathway to obesity and/or cancer?
% 	\vskip2ex
% 	Can we use it to narrow down our search for obesity-associated signature?
% \end{frame}

% \begin{frame}{Pathway Enrichment Analysis}
% 	Gene Ontology (GO) database was used to check for the function(s) being enriched by the DEGs in each cancer type (between obese and non-obese groups)
% 	\vskip2ex
% 	Used a rank-based approach and controlled the error rate using FDR
% 	\vskip2ex
% 	\begin{center}
% 		\textbf {No pathways showed up as significantly enriched}
% 	\end{center}
% \end{frame}

% \begin{frame}{Common genes and pathway enrichment analysis}
% 	There were no obesity-associated genes that were common across all the ICGC cancer types
% 	\vskip2ex
% 	There were no pathways enriched in any of the obesity-associated genes found so far
% \end{frame}

\begin{frame}{Relevance of BMI in Cris' data}
	\begin{center}
		\includegraphics[width=0.60\paperwidth]{./images/tmp/mikres}
	\end{center}
\end{frame}

\begin{frame}{So \ldots BMI is not related to cancer?}
	All the results so far suggests that there is no association between BMI and cancer
	\vskip2ex
	However, the metagenes are picking up gene signatures  that are common across all cancer types, but not related to BMI
	\vskip2ex
	So what signature/mechanism is the metagene picking up?
\end{frame}

\begin{frame}{Gatza \textit{et al.} (2010)}
	Used 18 pathways to define sub-class of breast tumours, based on the expression of the pathway signatures
	\begin{itemize}
		\item AKT, ER, PR, HER2, PI3K, p53, MYC, RAS, etc...
	\end{itemize}
	\vskip2ex
	Each pathway signature was obtained by overexpressing a single gene in an appropriate cell line, and then differentially expressed genes were noted down
	\begin{itemize}
			\item Use these pathways to see what the obesity-associated genes are most similar to
	\end{itemize}
\end{frame}

\begin{frame}{Tricky thing about metagene direction}
	Need to consider how the original data was manipulated:
	\begin{enumerate}
		\item Normalisation methods (RMA vs MAS5)
		\item Scaling the data or not
		\item Metagene scaling (probit vs rank-based)
	\end{enumerate}
	\vskip2ex
	Need to check the direction of each pathway metagenes with its gene (e.g. EGFR gene with EGFR pathway metagene)
	\vskip2ex
	Need to consider the grouping/clustering of the pathway metagenes
\end{frame}

\begin{frame}{What I mean about directions...}
	\includegraphics[width=0.45\paperwidth]{./images/output-0}
	\includegraphics[width=0.45\paperwidth]{./images/output-1}
\end{frame}

\begin{frame}{Some weird genes}
	\includegraphics[width=0.45\paperwidth]{./images/output-2}
	\includegraphics[width=0.45\paperwidth]{./images/output-3}
\end{frame}

\begin{frame}{Grouping/clustering of different pathways}
	\begin{center}
		\includegraphics[width=0.45\paperwidth]{./images/gatzares1}
	\end{center}
\end{frame}

\begin{frame}{Re-making results from Gatza \textit{et al.} paper}
	\includegraphics[width=0.45\paperwidth]{./images/out2}
	\includegraphics[width=0.40\paperwidth]{./images/gatzares1}
\end{frame}

\begin{frame}{Predicting obesity metagene with pathway metagene}
	Make a linear model (in Cris' data) with sample BMI and/or pathway metagenes and use this to predict Creighton's obesity metagene
	\vskip2ex
	\begin{itemize}
		\item If any of the pathway metagenes are significant in the model, the pathway is a good predictor of the obesity metagene
	\end{itemize}
	\vskip2ex
	Then use this model to predict the obesity metagene in Creighton data
	\begin{itemize}
		\item Compare the predicted obesity metagene with the SVD-generated obesity metagene
	\end{itemize}
\end{frame}

\begin{frame}{Results}
	Sample BMI was not significant  in the model
	\vskip2ex
	Only pathway that was a good predictor of obesity metagene was PR pathway metagene
\end{frame}

\begin{frame}{\normalsize Comparison of the predicted obesity metagene with true value}
	\begin{center}
		\includegraphics[width=0.45\paperwidth]{./images/tmp/prediction1}
		\includegraphics[width=0.45\paperwidth]{./images/tmp/prediction2}
	\end{center}
\end{frame}

\begin{frame}{\normalsize Comparison of the predicted obesity metagene with true value}
	\begin{center}
		\includegraphics[width=0.45\paperwidth]{./images/tmp/prediction3}
		\includegraphics[width=0.45\paperwidth]{./images/tmp/prediction4}
	\end{center}
\end{frame}

\begin{frame}{Overall Summary}
	\textbf{Creighton/FM metagene analysis}
	\begin{itemize}
		\item Both Creighton and FM metagenes correlated with the overall gene expression in any cancer data set
		\item The metagene score does not correlate with the BMI status of the samples
	\end{itemize}
	\vskip2ex
	\textbf{Gene expression and metagene analysis}
	\begin{itemize}
		\item Again, all of the gene sets found correlated with the overall gene expression in any data set, but none of them correlated with the BMI status
	\end{itemize}
\end{frame}

\begin{frame}{Overall Summary}
	\textbf{Gatza pathway metagene analysis}
	\begin{itemize}
		\item Sample BMI was not a good predictor of the obesity metagene, but PR pathway metagene was
		\item However, the model was not great at predicting the obesity metagene
			\begin{itemize}
				\item Other mechanism contributing to obesity metagene?
			\end{itemize}
	\end{itemize}
\end{frame}

% \begin{frame}{Discussion}
% 	Tumour heterogeneity within and between samples
% 	{\footnotesize
% 		\begin{itemize}
% 				% TODO rewrite the following sentence
% 			\item Not finding any significant genes or pathways because the samples are so different from one another
% 		\end{itemize}
% 	}
% 	\vskip2ex
% 	Tumour sample quality and size
% 	{\footnotesize
% 		\begin{itemize}
% 			\item Picking up too much noise and/or don't have enough samples to find anything significant (i.e. shit data)
% 		\end{itemize}
% 	}
% \end{frame}

% \begin{frame}{Discussion -- Inter-tumour heterogeneity 1}
% 	Each patient has an obesity driven pathway dysregulated, as well as different ``core'' pathways that lead to cancer progression
% 	\vskip2ex
% 	Each data set will have a different collection of patients, which in turn will have a different collection of dysregulated pathways
% 	{\footnotesize
% 		\begin{itemize}
% 			\item Hence DEGs show up, but not the pathways
% 		\end{itemize}
% 	}
% \end{frame}

% \begin{frame}{Discussion -- Inter-tumour heterogeneity 2}
% 	Some of the high BMI patients in the data set will have tumours that are BMI driven, but BMI may not be the driver for some of these patients
% 	\begin{itemize}
% 		\item Only subset of the patients with high BMI are actually driven by BMI
% 	\end{itemize}
% 	\vskip2ex
% 	Other data sets may have less of these truly BMI driven tumours
% 	\begin{itemize}
% 		\item Therefore no association with the signatures
% 	\end{itemize}
% \end{frame}

\begin{frame}{Acknowledgement}
	\begin{columns}
		\column{0.5\textwidth}
		My supervisor: Mik\\
		Members of the Black lab\\
		Cancer Genetics Lab\\
		SYSKA/Mozilla Study Group\\
		ICGC and TCGA\\
		\column{0.5\textwidth}
		\includegraphics[width=0.3\paperwidth]{./images/tcga}\\
		\includegraphics[width=0.3\paperwidth]{./images/icgc}
	\end{columns}
\end{frame}

% \begin{frame}
% 	\frametitle{References}
% 	\bibliographystyle{BiocRefStyle}
% 	\bibliography{/home/riku/Documents/References/BibTeX/MSc}
% \end{frame}

% \begin{frame}{test}
% 	\citet{Fuentes-Mattei2014}\\
% 	\citet{Creighton2012}\\
% 	\citet{Alexandrov2013}
% 	\cite{Gatza2010a}
% \end{frame}

\end{document}