### 2016-09-12 presentation/ctcr/beamer_theme/example.tex

\documentclass{beamer}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}

\usetheme{riku}

\def\logo{
	{\includegraphics[width=20mm]{logo}}
}

\title[BMI and Cancer Pathway]{Body Mass Index and Pathway Dysregulation in Cancer}
\subtitle{Committee Meeting 2016}
\author[Riku Takei]{Riku Takei (MSc)\\Supervisor: Assoc. Prof. Mik Black}
\institute{Department of Biochemistry, University of Otago}

\begin{document}

{
	\setbeamertemplate{footline}{}
	\begin{frame}[noframenumbering]
		\titlepage
	\end{frame}
}

\begin{frame}
	\frametitle{Introduction}
	blahabaldkfja
	\includegraphics{obesityworldmap.png}
\end{frame}

\end{document}