### 2016-10-22 thesis_plan/plan.tex
added thesis plan

\documentclass[a4paper, 10pt]{article}

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[margin=1.5cm]{geometry}

\begin{document}
	\section*{Introduction}
	\begin{itemize}
		\item BMI Intro.
		\item Creighton and Funtes-Mattei intro
		\item Gene signature intro + microarray/NGS intro
		\item BMI and cancer
		\item Mechanism of cancer progression in obese patient
		\item TCGA/ICGC/GEO
		\item Gatza pathway studies (old, 18 pathways and new papers)
		\item SVD/transformation matrix intro (maybe in methods?)
		\item Aim of research
	\end{itemize}
	\section*{Methods}
	\begin{itemize}
		\item WHO BMI classification
		\item Data download
		\item Normalisation methods (RMA/MAS5)
		\item Batch effect
		\item Standardising/scaling
		\item SVD/transformation matrix
		\item Gene expression analysis
		\item Pathway enrichment analysis
		\item Metagene direction
		\item Obesity metagene prediction with Gatza pathway (using linear model)
	\end{itemize}
	\section*{Results part 1 -- Creighton/FM metagene and Cris/ICGC data}
	\begin{itemize}
		\item Original Creighton obesity metagene in Creighton data
		\item Original Creighton obesity metagene in ICGC/Cris data
		\item Original FM obesity metagene in FM and Creighton data
		\item Original FM obesity metagene in ICGC/Cris data
		\item DEGs in Creighton data
		\item DEG obesity metagenes from Creighton data in ICGC/Cris data
		\item Common DEGs in ICGC data (per-cancer and combined)
		\item Path enrichment analysis in ICGC data (per-cancer and combined)
	\end{itemize}
	\section*{Results part 2 -- Gatza pathway metagenes and obesity metagenes}
	\begin{itemize}
		\item Gatza pathway metagene direction in Gatza data
		\item All metagenes in all data set, and correlation between SVD and transformation matrix
			\begin{itemize}
				\item Determining which data set the transformation matrices is being made in
			\end{itemize}
		\item Predicting obesity metagene (in Cris' data) using linear model that was made from Gatza pathways
	\end{itemize}
\end{document}