### 2017-04-02 notes/notebook.tex

% My master tips book
\documentclass[a4paper, 12pt] {report}

% All of the packages imported for this document:
\usepackage[utf8]{inputenc}
\usepackage{glossaries}
\usepackage{hyperref}
\usepackage{setspace}
\usepackage{url}
\usepackage{multicol}

\makeglossaries

\begin{document}

\include{./title/title}
\include{./glossary/glossary}

\tableofcontents

\printglossaries
\clearpage

\include{./subnotes/intro}

\include{./subnotes/gwas}

\end{document}