### 2017-04-02 notes/subnotes/intro.tex

%--------------------------------------------------------------------------------
%     File Name           :     subnotes/intro.tex
%     Created By          :     rikutakei
%     Creation Date       :     [2017-04-03 10:27]
%     Last Modified       :     [2017-04-03 10:37]
%     Description         :     Intro for my notebook
%--------------------------------------------------------------------------------

\chapter{Introduction}
\label{cha:introduction}

This notebook is for noting down all of the concepts and background information I need to know for my current and future projects.
I will try and maintain all of the references and keep this note as updated as possible.
\\

\noindent
This notebook is free of use for everyone, as long as it complies with the licence file associated with this Git repository.

\vfill

Riku Takei