### 2017-04-03 notes/notebook.tex
tmp commit

% My master tips book
\documentclass[a4paper, 12pt] {report}

% All of the packages imported for this document:
% \usepackage[utf8]{inputenc}
% \usepackage{glossaries}
% \usepackage{hyperref}
% \usepackage{setspace}
% \usepackage{url}
% \usepackage{multicol}

% Preamble:
\usepackage[T1]{fontenc}                                 % package for specifying output font encoding
\usepackage[acronym,style=super,nogroupskip,section=chapter]{glossaries} % package for making glossaries
\usepackage[flushleft]{threeparttablex}                  % package to add notes at the bottom of the table
\usepackage[font=footnotesize,format=hang]{caption}      % package for formatting figure caption
\usepackage[hidelinks, colorlinks]{hyperref}             % package for hyperlinking references - hidelinks option removes the ugly boxes around the huperlinks
\usepackage[md]{titlesec}                                % package for tweaking the chapter/section/subsection titles format
\usepackage[noabbrev,capitalise]{cleveref}               % package for referencing multiple referencing in a single \ref command
\usepackage[toc]{appendix}                               % package for making appendices
\usepackage[utf8]{inputenc}                              % package for specifying input font encoding
\usepackage{chngcntr}                                    % package for making continuous figure numbering, instead of section-based numbering
\usepackage{graphicx}                                    % package for graphics handling (i.e. figure images)
\usepackage{mathptmx}                                    % package for Times New Roman font
\usepackage{natbib}                                      % package for reference management
\usepackage{setspace}                                    % package for setting line space (used in title page)
\usepackage{xcolor,colortbl}                             % package for making colours
\usepackage{booktabs}                                    % package for better looking tables
\usepackage{longtable}                                   % package for tables that span multiple pages
\usepackage{siunitx}                                     % package for scientific notations and units
\usepackage{etoolbox}                                    % package to 'robustify' \bfseries
\usepackage{enumitem}                                    % package for tweaking lists

% Set up the hyperlinks - change the colour of the links to blue:
% \hypersetup{allcolors=blue}
\hypersetup{allcolors=black} % uncomment this line to make the hyperlinks black

\def\labelitemi{--}

\makeglossaries

\begin{document}

\include{./title/title}
\include{./glossary/glossary}

\tableofcontents

\printglossaries
\clearpage

\include{./subnotes/intro}

\include{./subnotes/gwas}

\bibliographystyle{BiocRefStyle}

\addcontentsline{toc}{chapter}{\bibname}

% Print out the bibliography:
\bibliography{./GWAS}

\end{document}