### 2018-10-09 notes/notebook.tex
tmp commit

\documentclass[a4paper, 12pt, oneside]{report} % for final draft compilation
% \documentclass[a4paper, 12pt, oneside, draft]{report} % for draft compilation

% Preamble:
\usepackage[T1]{fontenc}                                 % package for specifying output font encoding
\usepackage[acronym,style=super,nogroupskip]{glossaries} % package for making glossaries
\usepackage[flushleft]{threeparttablex}                  % package to add notes at the bottom of the table
\usepackage[font=footnotesize,format=hang]{caption}      % package for formatting figure caption
\usepackage[hidelinks, colorlinks]{hyperref}             % package for hyperlinking references - hidelinks option removes the ugly boxes around the huperlinks
\usepackage[md]{titlesec}                                % package for tweaking the chapter/section/subsection titles format
\usepackage[noabbrev,capitalise]{cleveref}               % package for referencing multiple referencing in a single \ref command
\usepackage[toc]{appendix}                               % package for making appendices
\usepackage[utf8]{inputenc}                              % package for specifying input font encoding
\usepackage{chngcntr}                                    % package for making continuous figure numbering, instead of section-based numbering
\usepackage{graphicx}                                    % package for graphics handling (i.e. figure images)
\usepackage{mathptmx}                                    % package for Times New Roman font
\usepackage{natbib}                                      % package for reference management
\usepackage{setspace}                                    % package for setting line space (used in title page)
\usepackage{xcolor,colortbl}                             % package for making colours
\usepackage{booktabs}                                    % package for better looking tables
\usepackage{longtable}                                   % package for tables that span multiple pages
\usepackage{siunitx}                                     % package for scientific notations and units
\usepackage{etoolbox}                                    % package to 'robustify' \bfseries
\usepackage{enumitem}                                    % package for tweaking lists

% Define colour(s) for table rows:
\definecolor{LightCyan}{rgb}{0.88,1,1}

% Change the itemize bullet points to dash:
\def\labelitemi{--}

% Set up the hyperlinks - change the colour of the links to blue:
% \hypersetup{allcolors=blue}
\hypersetup{allcolors=black} % uncomment this line to make the hyperlinks black

% Change the default cref name for equations:
\creflabelformat{equation}{#2\textup{#1}#3}

% Change the default cref name for appendix:
\crefname{appcha}{Appendix}{Appendices}
\crefname{appsec}{Appendix}{Appendices}

% Make siunitx detect bold/italics/etc:
\robustify\bfseries
\sisetup{detect-all}

% Set the width of captions for threeparttable to \linewidth:
\captionsetup[threeparttable]{width=\linewidth}

% Set up continuous figures:
\DeclareCaptionFormat{cont}{#1 (cont.)\par}

% Change the default url font:
\urlstyle{rm}

% change figure numbering from section-based to continuous:
\counterwithout{figure}{chapter}
\counterwithout{equation}{chapter}
\counterwithout{table}{chapter}

% Change the title of bibliography to references:
\renewcommand{\bibname}{References}

% Change the title of Table of Contents:
\renewcommand{\contentsname}{Table of Contents}

% Change the title of Table of Contents:
% \renewcommand{\thesubfigure}{\Alph{subfigure}}

% Italicise the subsubsection title:
\titleformat{\subsubsection}{\normalfont\itshape\bfseries}{\thesubsubsection}{1em}{}
\titleformat{\paragraph}{\normalfont\itshape}{\theparagraph}{\noindent}{}

% Remove unnecessary spaces between the figures from different chapters in LoF and LoT:
\newcommand*{\noaddvspace}{\renewcommand*{\addvspace}[1]{}}
\addtocontents{lof}{\protect\noaddvspace}
\addtocontents{lot}{\protect\noaddvspace}

% For generating glossaries:
\include{./glossary/glossary}
\makeglossaries

% Specify graphics path:
\graphicspath{{./images/}}
\DeclareGraphicsExtensions{.pdf,.png,.jpg}

% Define some alias for very long citations:
% \defcitealias{TCGA}{NCI and NHGRI, 2016}
% \defcitealias{ICGC}{ICGC, 2016}
% \defcitealias{GEO}{NCBI, 2016}

\begin{document}

\include{./title/title}

\pagenumbering{roman} % set page numbering to roman numeral

\include{./subnotes/preface}

\addcontentsline{toc}{section}{Preface}

% \include{./misc/acknowledgement}

\addcontentsline{toc}{section}{\contentsname}

\tableofcontents

\clearpage

% \addcontentsline{toc}{section}{\listfigurename}

% \listoffigures

% \clearpage

% \addcontentsline{toc}{section}{\listtablename}

% \listoftables

% \clearpage

\addcontentsline{toc}{section}{Glossary}

\printglossary[style=super]

% Need to set line spacing to single space, otherwise it inserts a blank page for an unknown reason:
% \singlespacing

\clearpage

\addcontentsline{toc}{section}{Acronyms}

\printglossary[type=\acronymtype,style=super]

\clearpage

\pagenumbering{arabic} % revert the page numbering to normal numbering
\glsresetall

\include{./subnotes/gwas}

% \include{./methods/methods}

% \include{./results/results}

% \include{./results/results2}

% \include{./discussion/discussion}

% \include{./misc/appendix}

% \include{./misc/appendix2}

% Use the reference style specified in BiocRefStyle.bst:
\bibliographystyle{BiocRefStyle}

\addcontentsline{toc}{chapter}{\bibname}

% Print out the bibliography:
\bibliography{./GWAS}

\end{document}