### 2018-10-09 coding/algorithms/algorithm.tex
started writing algorithm a week

\documentclass[a4paper, 12pt, oneside]{report} % for final draft compilation
% \documentclass[a4paper, 12pt, oneside, draft]{report} % for draft compilation

% Preamble:
\usepackage[T1]{fontenc}                                 % package for specifying output font encoding
\usepackage[acronym,style=super,nogroupskip]{glossaries} % package for making glossaries
\usepackage[hidelinks, colorlinks]{hyperref}             % package for hyperlinking references - hidelinks option removes the ugly boxes around the huperlinks
\usepackage[md]{titlesec}                                % package for tweaking the chapter/section/subsection titles format
\usepackage[utf8]{inputenc}                              % package for specifying input font encoding
\usepackage{mathptmx}                                    % package for Times New Roman font

% Change the itemize bullet points to dash:
\def\labelitemi{--}

% Set up the hyperlinks - change the colour of the links to blue:
% \hypersetup{allcolors=blue}
\hypersetup{allcolors=black} % uncomment this line to make the hyperlinks black

% Change the default url font:
\urlstyle{rm}

\begin{document}

\include{./titlepage}

\include{./preface}

\end{document}